---
output: html_document
---

# General parameters

The following are parameters used for both ABM and Compartmental models.

```{r global-params}
EPI_BETA  <- 0.75
EPI_GAMMA <- 0.33
EPI_LATENCY <- 1/0.33
EPI_N     <- 10000
EPI_0     <- 0.01
EPI_NDAYS <- 50

Sys.setenv(
    EPI_BETA  = EPI_BETA,
    EPI_GAMMA = EPI_GAMMA,
    EPI_LATENCY = EPI_LATENCY,
    EPI_N     = EPI_N,
    EPI_0     = EPI_0,
    EPI_NDAYS = EPI_NDAYS
)
```

# Compartmental Models

## SIR Model

```{r setup}
library(deSolve)
library(ggplot2)
library(data.table)

# Code from
# Chapter 2: SIR
# Book "Epidemics: Models and Data using R"
# By: Ottar N. Bjørnstad
sirmod <- function(t, y, parms) {

    # Pull state variables from y vector
    S = y[1]
    I = y[2]
    R = y[3]
    
    # Pull parameter values from parms vector
    beta  = parms["beta"]
    mu    = parms["mu"]
    gamma = parms["gamma"]
    N     = parms["N"]
    
    # Define equations
    dS  = mu * (N - S) - beta * S * I/N
    dI  = beta * S * I/N - (mu + gamma) * I
    dR  = gamma * I - mu * R
    res = c(dS, dI, dR)
    
    # Return list of gradients
    list(res)

}

# Initial parameters
times <- seq(0, EPI_NDAYS, by = 1)
parms <- c(mu = 0, N = EPI_N, beta = EPI_BETA, gamma = EPI_GAMMA)
start <- c(S = EPI_N * (1 - EPI_0), I = EPI_N * EPI_0, R = 0)
```

```{r run-model}
out <- ode(y = start, times = times, func = sirmod, parms= parms)
out <- as.data.frame(out)
out <- rbind(
    with(out, data.table(date = time, status = "susceptible", counts = S)),
    with(out, data.table(date = time, status = "infected", counts = I)),
    with(out, data.table(date = time, status = "recovered", counts = R))
)
```

Now we visualize the model

```{r visualize}
ggplot(out, aes(x = date, y = counts)) +
    geom_line(aes(colour = status)) +
    labs(title = "Compartmental SIR")
```

## SEIR Model

```{r setup-seir}
# Code adapted from
# Chapter 2: SIR
# Book "Epidemics: Models and Data using R"
# By: Ottar N. Bjørnstad
seirmod <- function(t, y, parms) {

    # Pull state variables from y vector
    S = y[1]
    E = y[2]
    I = y[3]
    R = y[4]
    
    # Pull parameter values from parms vector
    beta  = parms["beta"]
    mu    = parms["mu"]
    alpha = parms["alpha"]
    gamma = parms["gamma"]
    N     = parms["N"]
    
    # Define equations
    dS  = mu * (N - S) - beta * S * I/N - mu * S
    dE  = beta * S * I/N - (mu + alpha) * E
    dI  = alpha * E - (mu + gamma) * I
    dR  = gamma * I - mu * R
    res = c(dS, dE, dI, dR)
    
    # Return list of gradients
    list(res)

}

# Initial parameters
parms <- c(
    mu = 0, N = EPI_N, beta = EPI_BETA,
    alpha = 1/EPI_LATENCY, gamma = EPI_GAMMA
    )
start <- c(S = EPI_N * (1 - EPI_0), E = EPI_N * EPI_0, I = 0, R = 0)
```

```{r run-seir-model}
out_seir <- ode(y = start, times = times, func = seirmod, parms = parms)
out_seir <- as.data.frame(out_seir)
out_seir <- rbind(
    with(out_seir, data.table(date = time, status = "susceptible", counts = S)),
    with(out_seir, data.table(date = time, status = "exposed", counts = E)),
    with(out_seir, data.table(date = time, status = "infected", counts = I)),
    with(out_seir, data.table(date = time, status = "recovered", counts = R))
)
```

Now we visualize the model

```{r visualize-seir}
ggplot(out_seir, aes(x = date, y = counts)) +
    geom_line(aes(colour = status)) +
    labs(title = "Compartmental SEIR")
```

# Agent-Based Model Approach

Calculation of the expected number of days in state $a$ when prob of changing
state equals $\alpha$ is $1/\alpha$

```{r a-prob}
set.seed(712)
a <- .3
R <- matrix(runif(2e5 * 50), ncol = 50)
dat <- apply(R, 1, \(x) {
    which.max(x < a)
})

mean(dat) - 1 / a
```

## Mathematical preliminaries

That agent $i$ becomes infected can be computed as follow:

\renewcommand{\Pr}[1]{\text{P}\left(#1\right)}
\newcommand{\Prcond}[2]{\text{P}\left(#1\vphantom{#2}\right.\left|\vphantom{#1}#2\right)}
\newcommand{\E}[1]{\mathbb{E}\left(#1\right)}

\begin{equation}
\Pr{S_i\to I_i} = 1 - \Pr{\text{No infection}}
\end{equation}

At the same time, the probability of not becoming infected equals to the
probability of no infected agent transmiting the infection. The probability
that agent $j$ infects $i$ equals

\begin{equation}
\Pr{j\text{ infects }i} = %
    \overbrace{\Pr{j\text{ interacts with }i}}^{\beta}\times%
    \overbrace{\Prcond{\text{infection}}{j\text{ interacts with }i}}^{\text{Prob of transmission: }p_t}
\end{equation}

In this case, $\beta$ is parametrized such that its values are within $(0,1)$.
Since transmission from the $I$ infected agents happens independently, we finally have:

\begin{equation}
\Pr{S_i\to I_i} = 1 - \left[1 - \beta p_t\right]^I
\end{equation}

With the above equation, we can now calculate the change in the number of susceptible
agents. In this case, it equals to the expected number of new infections:

\begin{equation}
\E{\Delta S} = -S\times\left\{1 - \left[1 - \beta p_t\right]^I\right\}
\end{equation}

With the same parametrization in cannonical SIR model (Kermack and McKendrick), the
instantaneous change in the number of susceptible agent equals
$\frac{\delta}{\delta t}S = -S \beta I$. We can show
that, given $S$ and $I$, as $\beta\downarrow 0$, i.e., the population grows, both rates
converge to the same number. Formally

\begin{align}
& \lim_{\beta\downarrow 0} \frac{\E{\Delta S}}{\frac{\delta}{\delta t}S} \\
& \text{Following L'Hopital's rule} \\
& = \lim_{\beta\downarrow 0} \frac{\frac{\delta}{\delta \beta} \E{\Delta S}}{\frac{\delta^2}{\delta\beta\delta t}S} \\
& = \lim_{\beta\downarrow 0} \frac{-SI\left[1-\beta p_t\right]^{I-1}p_t}{-SI} \\
& \text{Since we assume }p_t = 1 \\
& = \lim_{\beta\downarrow 0} \left[1-\beta\right]^{I-1} \\
& = 1
\end{align}

The same can be shown for the change in the number recovered.

## Simulation study

Now, what happens with `epiworld`.

```{bash}
./09-sir-connected.o -n $EPI_N -b $EPI_BETA -d $EPI_NDAYS -p $EPI_0 -r $EPI_GAMMA -i 1.0 -s555599
```

```{r abm-sir}
library(ggplot2)
epiworld <- data.table::fread("total_hist.txt")
ggplot(epiworld, aes(x = date, y = counts)) +
    geom_line(aes(colour = status)) +
    labs(title = "ABM SIR")
```

```{bash}
./09-seir-connected.o -n $EPI_N -b $EPI_BETA -d $EPI_NDAYS -p $EPI_0 -r $EPI_GAMMA -i 1.0 -s555599 -l $EPI_LATENCY
```

```{r abm-seir}
library(ggplot2)
epiworld <- data.table::fread("total_hist.txt")
ggplot(epiworld, aes(x = date, y = counts)) +
    geom_line(aes(colour = status)) +
    labs(title = "ABM SEIR")
```



# Comparing ABM with Compartmental Models

To this end, we will compare the results of the first run of the Compartmental
model with 100 runs of the ABM, compute the confidence interval, and see how
likely is the compartmental model to fall within the trajectory of the ABM
simulation.

## SIR

```{bash}
./09-sir-connected.o -n $EPI_N -b $EPI_BETA -d $EPI_NDAYS -p $EPI_0 -r $EPI_GAMMA -i 1.0 -s555599 -e 100
```

```{r experiment100}
library(ggplot2)
library(data.table)
epiworld <- data.table::fread("09-sir-connected-experiments.csv")
epiworld <- epiworld[, .(
    min  = quantile(counts, probs = .025),
    mean = mean(counts),
    max  = quantile(counts, probs = .975)), by = .(date, status)]

# Merging Compartmental
epiworld <- merge(
    epiworld,
    out[, .(date = date, status = status, compartmental = counts)],
    by = c("date", "status")
)

setorder(epiworld, status, date)
ggplot(epiworld, aes(x = date, y = mean)) +
    geom_ribbon(aes(ymin = min, ymax = max, colour = status), alpha = 0.1) +
    geom_line(aes(x = date, y = compartmental, colour = sprintf("%s (compt)", status)))
```

It seems that, although both yield the same equilibria, compartmental models 
reach the highest point of the simulation earlier. This makes sense as within
a single day of the ABM simulation, compartmental models have more events 
taking place. Nonetheless, as predicted, as $\beta\downarrow 0$, the differences
become lesser. Furthermore, we could use the fact that the transition rates
are known to compute an adjustment.


## SEIR

```{bash}
./09-seir-connected.o -n $EPI_N -b $EPI_BETA -d $EPI_NDAYS -p $EPI_0 -r $EPI_GAMMA -i 1.0 -s555599 -e 100 -l $EPI_LATENCY
```

```{r experiment100-seir}
library(ggplot2)
library(data.table)
epiworld_seir <- data.table::fread("09-seir-connected-experiments.csv")
epiworld_seir <- epiworld_seir[, .(
    min  = quantile(counts, probs = .025),
    mean = mean(counts),
    max  = quantile(counts, probs = .975)), by = .(date, status)]

# Merging Compartmental
epiworld_seir <- merge(
    epiworld_seir,
    out_seir[, .(date = date, status = status, compartmental = counts)],
    by = c("date", "status")
)

setorder(epiworld_seir, status, date)
ggplot(epiworld_seir, aes(x = date, y = mean)) +
    geom_ribbon(aes(ymin = min, ymax = max, colour = status), alpha = 0.1) +
    geom_line(aes(x = date, y = compartmental, colour = sprintf("%s (compt)", status)))
```

## Rates

```{r comparing-rates-seir}
S <- 1000
rate_comp <- function(I,B) S * B * I

rate_abm <- function(I,B) S * (1 - (1 - B)^I)

op <- par(mfrow = c(3, 2))
for (i in c(1, 10, 100)) {
    curve(rate_comp(i, x), from = .01, to = 0.05)
    curve(rate_abm(i, x), from = .01, to = 0.05, add = FALSE, lty = 2)
}
par(op)
```
