---
output: html_document
---

```{r global-params}
EPI_BETA  <- 1.25
EPI_GAMMA <- 0.5
EPI_N     <- 10000
EPI_0     <- 0.01

Sys.setenv(
    EPI_BETA  = EPI_BETA,
    EPI_GAMMA = EPI_GAMMA,
    EPI_N     = EPI_N,
    EPI_0     = EPI_0
)
```

# Compartmental Models

```{r setup}
library(deSolve)
library(ggplot2)
library(data.table)

# Code from
# Chapter 2: SIR
# Book "Epidemics: Models and Data using R"
# By: Ottar N. Bjørnstad
sirmod <- function(t, y, parms) {

    # Pull state variables from y vector
    S = y[1]
    I = y[2]
    R = y[3]
    
    # Pull parameter values from parms vector
    beta  = parms["beta"]
    mu    = parms["mu"]
    gamma = parms["gamma"]
    N     = parms["N"]
    
    # Define equations
    dS  = mu * (N - S) - beta * S * I/N
    dI  = beta * S * I/N - (mu + gamma) * I
    dR  = gamma * I - mu * R
    res = c(dS, dI, dR)
    
    # Return list of gradients
    list(res)

}

# Initial parameters
times <- seq(0, 26, by = 1)
parms <- c(mu = 0, N = EPI_N, beta = EPI_BETA, gamma = EPI_GAMMA)
start <- c(S = EPI_N * (1 - EPI_0), I = EPI_N * EPI_0, R = 0)
```

```{r run-model}
out <- ode(y = start, times = times, func = sirmod, parms= parms)
out <- as.data.frame(out)
out <- rbind(
    with(out, data.table(date = time, status = "susceptible", counts = S)),
    with(out, data.table(date = time, status = "infected", counts = I)),
    with(out, data.table(date = time, status = "recovered", counts = R))
)
```

Now we visualize the model

```{r visualize}
ggplot(out, aes(x = date, y = counts)) +
    geom_line(aes(colour = status)) +
    labs(title = "Compartmental SIR")
```

# Agent-Based Model Approach

## Mathematical preliminaries

That agent $i$ becomes infected can be computed as follow:

\renewcommand{\Pr}[1]{\text{P}\left(#1\right)}
\newcommand{\Prcond}[2]{\text{P}\left(#1\vphantom{#2}\right.\left|\vphantom{#1}#2\right)}
\newcommand{\E}[1]{\mathbb{E}\left(#1\right)}

\begin{equation}
\Pr{S_i\to I_i} = 1 - \Pr{\text{No infection}}
\end{equation}

At the same time, the probability of not becoming infected equals to the
probability of no infected agent transmiting the infection. The probability
that agent $j$ infects $i$ equals

\begin{equation}
\Pr{j\text{ infects }i} = %
    \overbrace{\Pr{j\text{ interacts with }i}}^{\beta}\times%
    \overbrace{\Prcond{\text{infection}}{j\text{ interacts with }i}}^{\text{Prob of transmission: }p_t}
\end{equation}

In this case, $\beta$ is parametrized such that its values are within $(0,1)$.
Since transmission from the $I$ infected agents happens independently, we finally have:

\begin{equation}
\Pr{S_i\to I_i} = 1 - \left[1 - \beta p_t\right]^I
\end{equation}

With the above equation, we can now calculate the change in the number of susceptible
agents. In this case, it equals to the expected number of new infections:

\begin{equation}
\E{\Delta S} = -S\times\left\{1 - \left[1 - \beta p_t\right]^I\right\}
\end{equation}

With the same parametrization in cannonical SIR model (Kermack and McKendrick), the
instantaneous change in the number of susceptible agent equals
$\frac{\delta}{\delta t}S = -S \beta I$. We can show
that, given $S$ and $I$, as $\beta\downarrow 0$, i.e., the population grows, both rates
converge to the same number. Formally

\begin{align}
& \lim_{\beta\downarrow 0} \frac{\E{\Delta S}}{\frac{\delta}{\delta t}S} \\
& \text{Following L'Hopital's rule} \\
& = \lim_{\beta\downarrow 0} \frac{\frac{\delta}{\delta \beta} \E{\Delta S}}{\frac{\delta^2}{\delta\beta\delta t}S} \\
& = \lim_{\beta\downarrow 0} \frac{-SI\left[1-\beta p_t\right]^{I-1}p_t}{-SI} \\
& \text{Since we assume }p_t = 1 \\
& = \lim_{\beta\downarrow 0} \left[1-\beta\right]^{I-1} \\
& = 1
\end{align}

The same can be shown for the change in the number recovered. 

## Simulation study

Now, what happens with `epiworld`.

```{bash}
./09-sir-connected.o -n $EPI_N -b $EPI_BETA -d25 -p $EPI_0 -r $EPI_GAMMA -i 1.0 -s555599
```

```{r}
library(ggplot2)
epiworld <- data.table::fread("total_hist.txt")
ggplot(epiworld, aes(x = date, y = counts)) +
    geom_line(aes(colour = status)) +
    labs(title = "ABM SIR")
```

# Comparing ABC with Compartmental Models

To this end, we will compare the results of the first run of the Compartmental
model with 100 runs of the ABC, compute the confidence interval, and see how
likely is the compartmental model to fall within the trajectory of the ABC
simulation.

```{bash}
./09-sir-connected.o -n $EPI_N -b $EPI_BETA -d25 -p $EPI_0 -r $EPI_GAMMA -i 1.0 -s555599 -e 100
```

```{r experiment100}
library(ggplot2)
library(data.table)
epiworld <- data.table::fread("09-sir-connected-experiments.csv")
epiworld <- epiworld[, .(
    min  = quantile(counts, probs = .025),
    mean = mean(counts),
    max  = quantile(counts, probs = .975)), by = .(date, status)]

# Merging Compartmental
epiworld <- merge(
    epiworld,
    out[, .(date = date, status = status, compartmental = counts)],
    by = c("date", "status")
)

setorder(epiworld, status, date)
ggplot(epiworld, aes(x = date, y = mean)) +
    geom_ribbon(aes(ymin = min, ymax = max, colour = status), alpha = 0.1) +
    geom_point(aes(x = date, y = compartmental, colour = status))
```
