---
output: github_document
editor_options: 
  chunk_output_type: console
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# epiworld

<!-- badges: start -->
<!-- badges: end -->

The goal of epiworld is to ...

## Installation

You can install the development version of epiworld from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("gvegayon/world-epi")
```

## Example

This is a basic example which shows you how to solve a common problem:

```{r example}
library(epiworld)
library(igraph)

m <- new_epi_model()
geneseq <- c(T, F, T, T, F)

# Creating tools
add_tool_immune(m, !geneseq, 1)
add_tool_vaccine(m, !geneseq, .5)
add_tool_mask(m, !geneseq, .25)

# And the virus
add_virus_covid19(m, geneseq, .05)

# Adding people
set.seed(1231)
net   <- igraph::sample_smallworld(1, 1000, 4, .1)
edges <- igraph::as_edgelist(net)

edgelist_from_vec(m, edges[,1], edges[,2], directed = TRUE)

# Virus and tools are distributed
init_epi_model(m, 100, 123)

# We can get information about the model
m

# And run the model
run_epi_model(m)
m

```

Can run multiple times (in parallel):

```{r}
verbose_off_epi_model(m)

ans <- parallel::mclapply(1:1000, function(i) {
  
  # Running and resetting the model
  reset_epi_model(m)
  run_epi_model(m)
  
  # Retrieving the model information
  data.frame(
    replicate  = i,
    date       = get_hist_variant(m, "date"),
    id         = get_hist_variant(m, "id"),
    nrecovered = get_hist_variant(m, "date"),
    ninfected  = get_hist_variant(m, "ninfected"),
    ndeceased  = get_hist_variant(m, "ndeceased")
  )
  
}, mc.cores = 6L)

ans <- do.call(rbind, ans)
```

Visualizing

```{r}
library(ggplot2)
ggplot(ans[ans$id == 0,], aes(x = as.integer(date), y = ninfected)) +
  geom_point(alpha = .5, color = "grey") +
  labs(
    x = "Date", y = "Infected with COVID",
    caption = "Includes 1000 replicates")
```


## Code of Conduct

Please note that the epiworld project is released with a [Contributor Code of Conduct](https://contributor-covenant.org/version/2/0/CODE_OF_CONDUCT.html). By contributing to this project, you agree to abide by its terms.